---
title: "Is it all over?"
author: Robin Penfold
date: 2019-05-07
categories:
  - prempredict
  - simulation
draft: false
preview: PPLOGO.jpeg
slug: they-think-it-is-all-over
description: |
  I run 10,000 simulations on the final round of matches in the Premier League season, so as to project the odds of each player winning my PremPredict competition.
output:
  distill::distill_article:
    self_contained: false
    highlight: pygments
---

> "Some people are on the pitch. They think it's all over ..." 

But it isn't. It could be Roger or Luke.

## Current standings

Let's start with the current standings:

```{r ttiao1, eval=TRUE, echo=FALSE, message=FALSE, warning=FALSE, results='hide'}
library(premPredictor)
suppressMessages(library(tidyverse))

iterations <- 10000
  
player_info <- 
  "https://www.dropbox.com/s/uin6zk4w5cyk2m1/PremPredict-18-19.csv"
player_data <- get_player_data(url_value = player_info)

odds_info <- "https://www.dropbox.com/s/cb5qyxyexksvez1/odds.csv"
odds_data <- get_odds_data(url_value = odds_info)

results <- get_latest_standings(
  data_input = player_data, 
  use_saved_data = T, 
  data_file = '201819_penultimate.rds'
  )
```

```{r ttiao1a, echo=FALSE, layout="l-body"}
knitr::kable(results)
```

## Who will win PremPredict?

But here's what we really want to know: What are our chances of winning? 

```{r ttiao2, eval=TRUE, echo=FALSE, message=FALSE, warning=FALSE}
project_players <- get_projection(
  url_odds = odds_info,
  url_predictions = player_info,
  project_players = TRUE,
  use_saved_data = T,
  data_file = "201819_penultimate.rds",
  runs = iterations)

winPlayer <- as.data.frame(project_players==1L)

winLikelihood <- as.data.frame(
  apply(winPlayer, 2, sum)
)

winSummary <- tibble(
  Name = colnames(project_players), 
  Frequency = pull(winLikelihood, 1)) %>% 
  filter(Frequency > 0) %>% 
  arrange(-Frequency) %>%
  mutate(Likelihood = Frequency/iterations) %>%
  select(Name, Likelihood)

knitr::kable(winSummary)
```

Well, unless your name is Roger or Luke then the answer is zero. (Or, more specifically, in the 10,000 scenarios that I ran, you didn't win any of them.)

## Who will win the Premier League?

By the way, these calculations also show who the bookies feel will be champions. 

```{r ttiao3, eval=TRUE, echo=FALSE, message=FALSE, warning=FALSE}
project_team <- get_projection(
  url_odds = odds_info, 
  url_predictions = player_info, 
  project_players = FALSE,
  use_saved_data = T,
  data_file = "201819_penultimate.rds",
  runs = iterations)

project_team %>% 
  select(Liverpool, `Man City`) %>% 
  count(Liverpool, `Man City`) %>% 
  mutate(chance = n / sum(n)) %>% 
  arrange(desc(chance)) %>% 
  knitr::kable()
```

## Does each player depend upon a team to win?

In short, will Luke really support Liverpool?

```{r ttiao4, eval=TRUE, echo=FALSE, message=FALSE, warning=FALSE}
bind_cols(winPlayer, project_team) %>% 
  select(
    Roger = Roger.Gathercole, 
    Luke = Luke.Finnis, 
    `Man City`, Liverpool) %>% 
  count(Roger, Luke, `Man City`, Liverpool) %>% 
  mutate(chance = n / sum(n)) %>% 
  arrange(`Man City`, desc(chance)) %>% 
  knitr::kable()
```

##   {.appendix}

For completeness (and reproducibility), here's the code that I used to calculate what's above.

## Current standings {.appendix}

```{r ttiao1, eval=FALSE, echo=TRUE}
```

```{r ttiao1a, eval=FALSE, echo=TRUE}
```

## Who will win PremPredict? {.appendix}

```{r ttiao2, eval=FALSE, echo=TRUE}
```

## Who will win the Premier League? {.appendix}

```{r ttiao3, eval=FALSE, echo=TRUE}
```

## Does each player depend upon a team to win? {.appendix}

```{r ttiao3, eval=FALSE, echo=TRUE}
```

## System settings {.appendix}

```{r, eval=TRUE, echo=FALSE}
sessionInfo()
```
