---
title: "Imputing tracking error from active share and the number of stocks"
author: Robin Penfold
date: 2018-08-09
categories:
  - papers
  - performance
draft: false
slug: tracking-error-alpha-share
preview: tracking-error2.jpg
bibliography: Sapra.bib
description: |
  Some  done.
output:
  distill::distill_article:
    self_contained: false
    highlight: pygments
---

```{r te_as1, include=FALSE}
knitr::opts_chunk$set(echo = FALSE)
```

Let's start with some theory.

## Theory

As Steve Sapra at PIMCO shows [@Sapra2013], tracking error can be linearly expressed in terms of active share ...

After a little manipulation, it can be shown that:

\(S = \omega / (\bar{\sigma _{e}} \sqrt{2\pi / n})\)    , where:

  * \(S\) -- the expected active share of the product
  
  * \(\omega\) -- the expected tracking error of the product
  
  * \(\bar{\sigma _{e}}\) -- the average idiosyncratic risk of all stocks in the investible universe
  
  * \(n\) -- the number of active stocks in the product

  
Put another way:

\(\omega / S = \bar{\sigma _{e}} \sqrt{2\pi / n} = \bar{\sigma _{e}} \sqrt{2\pi}(1 / \sqrt{n})\)

In other words, if I were to regress \(\omega / S\) against \(1 / \sqrt{n}\) for all the products in a mandate, the resulting beta coefficient would be the empirical estimate of \(\bar{\sigma _{e}} \sqrt{2\pi}\) -- which I could divide by \(\sqrt{2\pi}\) to obtain the emprical estimate of \(\bar{\sigma _{e}}\) for the mandate.

etc 

***

For completeness and reproducibility, I now show the calculations used in this analysis.

## Calibration {.appendix}

```{r te_as2, eval=FALSE, echo=TRUE}
```

</br>

## Coefficients {.appendix}

```{r te_as3, eval=FALSE, echo=TRUE}
```

</br>

## Conclusion {.appendix}

```{r te_as4, eval=FALSE, echo=TRUE}
```

</br>

## System settings {.appendix}

```{r, eval=TRUE, echo=FALSE}
sessionInfo()
```
