---
title: "PremPredict 2019/20"
author: Robin Penfold
date: 2019-08-10
categories:
  - prempredict
draft: false
slug: twenty-one-teams
description: |
  In this year's PremPredict, there are twenty teams, twenty-one players and a massive £105 up for grabs. In this post, I'll show the latest standings and our collective expectations for the season.
output:
  distill::distill_article:
    self_contained: false
    highlight: pygments
---

Welcome to the 2019/20 season of PremPredict!

Rather than create new posts every few months, I thought you'd appreciate a page that updated with the latest results every time that you hit refresh.

With that in mind, here are the very latest standings.

## Latest standings

```{r thecode201, echo=FALSE, results='hide', error=TRUE, message=FALSE, warning=FALSE, layout="l-body-outset"}
library(ggridges)
library(premPredictor)
suppressMessages(library(tidyverse))
  
# player_info <- 
#   "https://www.dropbox.com/s/uin6zk4w5cyk2m1/PremPredict-19-20.csv"
# player_data <- get_player_data(url_value = player_info)

player_data <- read_rds('data_input_2019-20.rds')

results <- get_latest_standings(
  data_input = player_data, 
  use_saved_data = T, 
  data_file = '201920_week1.rds'
  )
```

<iframe width="800" height="417" scrolling="yes" frameborder="no"  src="https://robin.shinyapps.io/PP-shiny/"> </iframe>

## Collective expectations

But what are we collectively expecting from the Premier League this season? Are we expecting it to be a repeat of last season?

By the look of our picks, we seem to side with the bookies. As a group, we predict Manchester City to finish highest on average, with the recently-promoted teams struggling.

```{r collExp20, echo=FALSE, error=TRUE, layout="l-body-outset", preview=TRUE}
data_input1 <- as_tibble(player_data)
averageView <- round(rowMeans(data_input1[, -1]), 2)
views <- cbind(data_input1[,1], averageView)
  
data_input2 <- data_input1 %>% 
  gather(key = "Player", -Club, value = "Prediction") %>% 
  left_join(views, by = "Club")
  
ggplot(
  data = data_input2,
  mapping = aes(
    y = reorder(Club, -averageView), 
    x = Prediction, 
    fill = averageView, 
    color = averageView
    )
  ) + 
  geom_ridgeline(
    stat = "binline", 
    bins = 20, scale = 0.95, 
    draw_baseline = FALSE
    ) + 
  scale_x_continuous(
    breaks = c(5, 10, 15, 20), 
    labels = c(5, 10, 15, 20)
    ) + 
  labs(
    y = "", x = "", 
    title = "\n Our collective predictions for this season \n"
    ) + 
  scale_fill_gradient(
    low = "green", high = "red", 
    guide=FALSE
    ) + 
  scale_color_gradient(
    low = "green", high = "red", 
    guide=FALSE
    ) +
  theme(
    title = element_text(size = 10), 
    axis.text.y = element_text(size = 6)
    )
```

And we collectively feel that Sheffield United will come last. But will we be more accurate than before?!

##   {.appendix}

For completeness (and reproducibility), here's the code that I used to calculate what's above. Note that the app does things a little differently, but this standalone code should give you the gist of the approach.

## Latest standings {.appendix}

```{r showthecode201, ref.label='thecode201', echo=TRUE, eval=FALSE}
```

## Collective expectations {.appendix}
  
```{r showcollExp20, ref.label='collExp20', echo=TRUE, eval=FALSE}
```

## Calculation inputs {.appendix}

And here are the two inputs to these calculations:

  * Your [team-by-team predictions](https://www.dropbox.com/s/uin6zk4w5cyk2m1/PremPredict-19-20.csv)
  
  * The [latest team standings in the Premier League](http://www.theguardian.com/football/premierleague/table)

## System settings {.appendix}

```{r, eval=TRUE, echo=FALSE}
sessionInfo()
```

## Wall of Fame {.appendix}

And here are our previous champions, including a double winner that is sorely missed.

  * 2007/08 -- Robert Wye
  * 2008/09 -- Les Penfold
  * 2009/10 -- Beth Penfold
  * 2010/11 -- George Quin and Miranda Stride
  * 2011/12 -- Hannah Finnis
  * 2012/13 -- Michael Cheeseman
  * 2013/14 -- Danny Russell
  * 2014/15 -- Mathew Saunders
  * 2015/16 -- Roger Gathercole
  * 2016/17 -- Luke Finnis
  * 2017/18 -- George Quin
  * 2018/19 -- Roger Gathercole
