---
title: "Noisy boys ahead"
author: Robin Penfold
date: 2019-08-10
categories:
  - prempredict
draft: false
slug: twenty-one-teams
description: |
  Welcome to this year's PremPredict, where there are twenty teams, twenty-one players and a massive £105 up for grabs. But what's worse is that the noisy boys are on top: Yes, the old man, the brother and the best man. Please make it change!
  In this post, I'll also consider our collective expectations for the season.
output:
  distill::distill_article:
    self_contained: false
    highlight: pygments
---

Welcome to the 2019/20 season of PremPredict!

And, after a whole week, let's start with where everyone is currently placed.

<p style="height: .5em;">&nbsp;</p>

## Latest standings

```{r thecode201, echo=FALSE, results='hide', error=TRUE, message=FALSE, warning=FALSE, layout="l-body-outset"}
library(ggridges)
library(premPredictor)
suppressMessages(library(tidyverse))
  
# player_info <- 
#   "https://www.dropbox.com/s/uin6zk4w5cyk2m1/PremPredict-19-20.csv"
# player_data <- get_player_data(url_value = player_info)

player_data <- read_rds('data_input_2019-20.rds')

results <- get_latest_standings(
  data_input = player_data, 
  use_saved_data = T, 
  data_file = '201920_week1.rds'
  )
```


```{r thecode201a, echo=FALSE, layout="l-body-outset"}
knitr::kable(results)
```

</br>

Fortunately, there's still a long way to go!

</br>

## Collective expectations

But what are we collectively expecting from the Premier League this season? Are we expecting it to be a repeat of last season?

By the look of our picks, we seem to side with the bookies. As a group, we predict Manchester City to finish highest on average, with the recently-promoted teams struggling.

```{r collExp20, echo=FALSE, error=TRUE, layout="l-body-outset", preview=TRUE}
data_input1 <- as_tibble(player_data)
averageView <- round(rowMeans(data_input1[, -1]), 2)
views <- cbind(data_input1[,1], averageView)
  
data_input2 <- data_input1 %>% 
  gather(key = "Player", -Club, value = "Prediction") %>% 
  left_join(views, by = "Club")
  
ggplot(
  data = data_input2,
  mapping = aes(
    y = reorder(Club, -averageView), 
    x = Prediction, 
    fill = averageView, 
    color = averageView
    )
  ) + 
  geom_ridgeline(
    stat = "binline", 
    bins = 20, scale = 0.95, 
    draw_baseline = FALSE
    ) + 
  scale_x_continuous(
    breaks = c(5, 10, 15, 20), 
    labels = c(5, 10, 15, 20)
    ) + 
  labs(
    y = "", x = "", 
    title = "\n Our collective predictions for this season \n"
    ) + 
  scale_fill_gradient(
    low = "green", high = "red", 
    guide=FALSE
    ) + 
  scale_color_gradient(
    low = "green", high = "red", 
    guide=FALSE
    ) +
  theme(
    title = element_text(size = 10), 
    axis.text.y = element_text(size = 6)
    )
```

<p>&nbsp;</p>

And we collectively feel that Sheffield United will come last. But will we be more accurate than before?!

***

For completeness (and reproducibility), here's the code that I used to calculate what's above.

## Latest standings {.appendix}

```{r showthecode201, ref.label='thecode201', echo=TRUE, eval=FALSE}
```

</br>

## Collective expectations {.appendix}
  
```{r showcollExp20, ref.label='collExp20', echo=TRUE, eval=FALSE}
```

## Calculation inputs {.appendix}

And here are the two inputs to these calculations:

  * Your [team-by-team predictions](https://www.dropbox.com/s/uin6zk4w5cyk2m1/PremPredict-19-20.csv)
  
  * The [latest team standings in the Premier League](http://www.theguardian.com/football/premierleague/table)

</br>

## System settings {.appendix}

```{r, eval=TRUE, echo=FALSE}
sessionInfo()
```
